---
title: "Specifying Preprocessing Recipe and Evaluation Metrics"
author: "Paul Johnson"
date: '`r Sys.Date()`'
output: 
  html_document: 
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  warning = FALSE,
  message = FALSE
  )

# import packages
library(dplyr)
library(recipes)
library(yardstick)

# import train/test data, resamples, recipe, and metrics
load(here::here("data", "vb_data.RData"))

```

We can use everything we've learned from the EDA to specify a preprocessing recipe that processes the data ready for the model training.

# Preprocessing Recipe

Having carried out some exploration of the data in the [previous notebook](./src/eda.Rmd) the following features were identified for the model:

-   Skill Type
-   Skill Subtype
-   Blockers
-   Team/Opponent Setter Position (Grouped 1-3 and 4-6)
-   Start and End X/Y Coordinates (plus interactions)

```{r recipe}

model_rec <- 
  # model formula
  recipe(
    evaluation_code ~
      skill_type + skill_subtype + blockers +
      team_setter_position + opponent_setter_position +
      start_coordinate_x + start_coordinate_y +
      end_coordinate_x + end_coordinate_y,
    data = train_df
  ) %>%
  # convert team setter position to 2HP and 3HP groups
  step_mutate(
    team_setter_position =
      case_when(
        team_setter_position %in% c(1, 6, 5) ~ "3HP",
        team_setter_position %in% c(2, 3, 4) ~ "2HP"
      ),
    # convert opponent setter position to 2HP and 3HP groups
    opponent_setter_position =
      case_when(
        opponent_setter_position %in% c(1, 6, 5) ~ "3HP",
        opponent_setter_position %in% c(2, 3, 4) ~ "2HP"
      )
  ) %>%
  # convert setter position variables to factors
  step_string2factor(
    team_setter_position,
    opponent_setter_position
  ) %>%
  # combine low frequency factor levels
  step_other(all_nominal(), threshold = 0.05) %>%
  # convert categorical variables to factors
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  # specify interactions between variables
  step_interact(
    terms = ~ end_coordinate_x:end_coordinate_y +
      start_coordinate_x:start_coordinate_y) %>%
  # remove variables that are heavily correlated with others
  step_corr(threshold = 0.8) %>%
  # remove no variance predictors which provide no predictive information
  step_nzv(all_predictors())
  
# store the baked recipe df
baked_df <-
  model_rec %>%
  prep() %>%
  bake(new_data=NULL)

```

Comments:

-   I don't think we need to normalize the numeric variables because all are between 0 - 25, and I think these are suitably similar scales.

We can have a look at the data after all the preprocessing steps have been carried out:

```{r glimpse}

# check everything looks good
glimpse(baked_df)

```

Everything looks good!

# Specify Evaluation Metrics

We can also set the evaluation metrics for the training process. We will use ROC-AUC as the primary evaluation metric but also checking on the Accuracy, F1-Score, and J-Index.

```{r metrics}

eval_metrics <-
  metric_set(roc_auc, accuracy, f_meas, j_index)

```

This isn't really a preprocessing step but the same evaluation metrics are reused throughout the remaining notebooks, so it makes sense to set the metrics now and save on code elsewhere.

# Save Recipe and Metrics

We will save the preprocessing recipe and metric set so that they can be imported in the notebooks that follow.

```{r save}

save(
  # recipe
  model_rec,
  # metrics
  eval_metrics,
  file = here::here(
    "data", "preprocessed.RData"
  )
)

```
